---
title: "Rstan"
format: html
editor: visual
---

## [RStan Getting Started](https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started) 

### Remove previous installations of rstan

```{r}
#| eval: false  
#| echo: true
 
# remove.packages(c("StanHeaders", "rstan"))
# if (file.exists(".RData")) file.remove(".RData")

```
### Configuring C++ Toolchain 

Prior to installing RStan, you need to configure your R installation to be able to compile C++ code. Follow the link below for your respective operating system for more instructions:

* [RTools42 Installation](https://cran.r-project.org/bin/windows/Rtools/rtools42/rtools.html)  
* [RTools42 Installer - click here](https://cran.r-project.org/bin/windows/Rtools/rtools42/files/rtools42-5355-5357.exe)  
* Path:  `C:\ProgramData\Microsoft\Windows\Start Menu\Program\Rtools 4.2\`  
* Path2:  `C:\Users\Robert\AppData\Roaming\Microsoft\Windows\Recent`  

* RTools4 Installation  


* install.packages("StanHeaders", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
install.packages("rstan", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))

* Verify that Rtools is being found:  
`Sys.which("make")`

## "C:\\rtools40\\usr\\bin\\make.exe" 

### Install packages from source

install.packages("jsonlite", type = "source")

### Verify c++ compiler  

* From Terminal run:  g++ --version  



### Install the preview of `rstan` 2.26 using:  

```{r}
# install.packages("StanHeaders", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
# install.packages("rstan", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))


```


### Latest Development Version 2.26X (unrelased)

```{r}

#| eval: false
#| echo: true 


# run the next line if you already have rstan installed
# remove.packages(c("StanHeaders", "rstan"))

# install.packages("rstan", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))


```

The downloaded binary packages are in
	`C:\Users\Robert\AppData\Local\Temp\Rtmp6joOih\downloaded_packages`

### Verify RStan 

```{r}
example(stan_model, package = "rstan", run.dontrun = TRUE)
```
