---
title: "brms installation"
format: html
editor: visual
---


After encountering issues trying to install rstan and brms, I entered into a conversation with ChatGPT to overcome the installation issues.

I am able to run the brms package.  There could remain an issue with CmdStan. Intallation of CmdStan and Cygwyn;

RTools is what gives you access to c++ compiler.  You may have to configure where these files are.
Mark Andrews:  "It is quite tricky."  "You have to just persevere."  





## Question:

Using Rstudio and R, I have installed '*cmdstanr*'.

Code: install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos"))) .

```{r}
#| echo: true 
#| eval: false
install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos"))) 
```

When I run the following code with \*brm( )\* , I get an error: Error: CmdStan path has not been set yet. See ?set_cmdstan_path.

I do not find cmdstan on my computer. How do I verify if cmdstan has been installed?

```{r}
#| echo: true
#| eval: false 
#| 
fit_complete_pool = brm(log_radon ~ floor, data=mn_radon)
summary(fit_complete_pool)
```

To verify if cmdstan has been installed on your computer, you can follow these steps:

1.  Open a new R session in RStudio.\
2.  Type `library(cmdstanr)` in the R console and press Enter.\
3.  If cmdstan has been installed on your computer, the library will load without any errors. If it has not been installed, you will get an error message saying that the package could not be found.\
4.  If cmdstan has not been installed, you can install it using the following code: **cmdstanr::install_cmdstan()**. This will install the latest version of cmdstan on your computer.
5.  Once cmdstan has been installed, you need to set the CmdStan path using **set_cmdstan_path()** function in R. This function takes the path to the CmdStan installation directory as an argument. For example, if you installed cmdstan in the default location, you can set the path using the following code: set_cmdstan_path("\~/cmdstan-2.28.1").\
6.  After setting the CmdStan path, you should be able to run the brm() function without any errors.

When you install cmdstanr in RStudio, *it does not automatically install cmdstan on your computer*. **You need to separately install cmdstan on your computer and set the CmdStan path in RStudio.**\
\#### To find the CmdStan path on your computer, you can follow these steps:

1.  First, you need to install cmdstan on your computer. You can download the latest version of cmdstan from the official website: https://mc-stan.org/users/interfaces/cmdstan. Choose the appropriate version for your operating system and follow the installation instructions.

2.  Once you have installed cmdstan on your computer, you need to find the directory where it is installed. The exact location of the directory will depend on where you installed cmdstan and your operating system. By default, the directory is usually named "cmdstan-{version}" and is located in the home directory.

3.  To find the CmdStan path, you can open a new R session in RStudio and use the set_cmdstan_path() function to set the path to the directory where cmdstan is installed. For example, if cmdstan is installed in the default location, you can set the path using the following code:

`**set_cmdstan_path("~/cmdstan-{version}")**`

4.  Replace "{version}" with the version number of cmdstan that you installed.

5.  Once you have set the CmdStan path, you should be able to use cmdstanr to run Bayesian models in RStudio without any issues.

`C:\Users\Robert\AppData\Local\R\win-library\4.2\cmdstanr\R\cmdstanr`, is the path to the cmdstanr package directory in R on a Windows system. It is not the path to the CmdStan installation directory.

To find the path to the CmdStan installation directory on your computer, you need to locate where you installed CmdStan. By default, CmdStan is installed in the home directory, and the directory is named "cmdstan-{version}".

Once you have located the CmdStan installation directory, you can set the CmdStan path in R using the set_cmdstan_path() function. For example, if CmdStan is installed in the default location on a Windows system, you can set the path using the following code:

set_cmdstan_path("C:/Users/Robert/cmdstan-{version}")

Replace "{version}" with the version number of CmdStan that you installed.

### CmdStan Installation Directory

The "installation directory" refers to the directory where you installed the CmdStan software, not the directory where you installed the cmdstanr R package.

When you install cmdstanr using the **install.packages("cmdstanr")** command in RStudio, *it only installs the R package, not the CmdStan software itself*.

**You need to separately download and install CmdStan on your computer**.

The CmdStan installation directory is the folder on your computer where you have installed CmdStan. The directory contains all the necessary files and subfolders for running the CmdStan software. By default, when you install CmdStan, it creates a folder named **"cmdstan-{version}"** in your home directory.

For example, on a Windows system, the default installation directory for CmdStan might be:

**C:/Users/YourUserName/cmdstan-{version}** where {version} is the version number of the CmdStan software that you have installed.

### check cmdstan toolchain()

Please make sure that you have installed CmdStan on your computer and that you have set the CmdStan path to the correct directory where CmdStan is installed.

You can try running the following code to check if the CmdStan path has been set correctly:

`**check_cmdstan_toolchain()**`

This will test if the CmdStan path has been set correctly and will give you feedback on the status of the CmdStan toolchain.

If the CmdStan path is still not set correctly, you can try resetting it using the `**set_cmdstan_path()**` function again, making sure to set the correct path to the directory where CmdStan is installed.

Once the CmdStan path is set correctly, you should be able to run the brm() function without any errors.

### Rtools42 is required to run *CmdStan*

The error message you received indicates that Rtools42 is required to run CmdStan with R version 4.2.2. Rtools is a set of tools that allow you to build and compile packages in R on Windows systems.

To fix this error, you need to install Rtools42 on your computer.

1.  You can download and install Rtools42 from the following website: [Download Rtools here](https://cran.r-project.org/bin/windows/Rtools/).

-   Make sure to download the version of Rtools that corresponds to the version of R you are using (in this case, R version 4.2.2).

2.  **Check cmdstan toolchain**

After installing Rtools42, you need to run the `**check_cmdstan_toolchain()**` function in R to make sure that everything is set up correctly. This function checks that the necessary toolchain components are installed and working correctly.

-   Run the following command in R to check the toolchain:

-   `**cmdstanr::check_cmdstan_toolchain()**`

This will test if the toolchain is set up correctly and will give you feedback on the status of the CmdStan toolchain.

If everything is set up correctly, you should be able to run the brm() function without any errors.

-   Attempting to set_cmdstan_path, I ran the following code which generated an error: `set_cmdstan_path("C:/Users/Robert/AppData/Local/R/win-library/4.2/cmdstanr/R/")`

-   Warning message:\

-   Can't find CmdStan makefile to detect version number. Path may not point to valid installation.

This error message indicates that RStudio could not find the\* **CmdStan makefile** to detect the version number. This may indicate that the CmdStan path you specified may not be pointing to a valid installation of CmdStan.

To resolve this error, you should ensure that the path you specified is correct and points to the correct directory where CmdStan is installed. You can try specifying the path to the directory where the CmdStan binary files are located, rather than the R package directory.

For example, if you have installed CmdStan in the default directory on a Windows system, you can set the path using the following command:

-   `**set_cmdstan_path("C:/Users/Robert/cmdstan-{version}")**`

Replace Robert with your Windows username and {version} with the version number of the CmdStan software that you have installed.

If you have already set the correct CmdStan path, try running the check_cmdstan_toolchain() function to see if everything is set up correctly. This function will check that the necessary toolchain components are installed and working correctly.

If you are still having trouble, you may need to reinstall CmdStan and ensure that it is installed correctly on your computer before attempting to set the CmdStan path again.

The version of cmdstanr is 0.5.3 rewrite the set_cmdstan_path with the correct version.

Sure, you can set the correct cmdstanr path in R by specifying the correct version number. Here is an example of how to do this:

-   `**set_cmdstan_path("C:/Users/Robert/cmdstan-{version}")**`

-   Replace Robert with your Windows username and {version} with the version number of the CmdStan software that you have installed. Since you mentioned that the version of cmdstanr that you are using is 0.5.3, the correct path should look like this:

-   `**set_cmdstan_path("C:/Users/Robert/cmdstan-2.28.1")**`

Make sure to replace Robert with your Windows username and use the correct version number of CmdStan that you have installed. Once you have set the correct path, you should be able to run Bayesian models using **brm()** in RStudio without any issues.

If the directory `C:/Users/Robert/cmdstan-2.28.1` does not exist on your computer, it is possible that CmdStan was not installed correctly or was installed in a different location.

You can try **reinstalling CmdStan** and making sure that it is installed in the correct location. By default, CmdStan installs in the home directory of your computer and creates a folder named "cmdstan-{version}". You should check that this folder exists and contains the necessary files.

Once you have located the CmdStan installation directory, you can set the CmdStan path in RStudio using the **set_cmdstan_path()** function. Make sure to use the correct path to the CmdStan installation directory. For example, if CmdStan is installed in the default location on a Windows system, you can set the path using the following code:

-   `**set_cmdstan_path("C:/Users/Robert/cmdstan-{version}")**`

Replace Robert with your Windows username and {version} with the version number of the CmdStan software that you have installed.

If you are still having trouble locating the CmdStan installation directory, you can try running the find_cmdstan() function in RStudio. This function will search your computer for the CmdStan installation directory and return the path if it is found.

Once you have set the CmdStan path correctly, you should be able to run the brm() function without any issues.

### Error:

set_cmdstan_path("C:/Users/Robert/AppData/Local/R/win-library/4.2/cmdstanr/R/cmdstanr") Which is the correct folder. But it returns the error: \* **Error: CmdStan path has not been set yet**. See ?set_cmdstan_path.

### I do not understand your instructions.

-   I can browse to the folder C:/Users/Robert/AppData/Local/R/win-library/4.2/cmdstanr/R/cmdstanr. \* That is the only folder with cmdstanr on my computer. The folder does not have a version number associated with it. Please explain.

I apologize for the confusion.\* **It seems that I made a mistake in my previous response**.

**If you have installed cmdstanr using the install.packages("cmdstanr") command in RStudio, then you do not need to specify a path to the CmdStan installation directory. cmdstanr will automatically locate the CmdStan installation directory for you.**

However, in order for cmdstanr to find the correct CmdStan installation directory,**you need to have installed CmdStan separately on your computer before installing cmdstanr**.

The error message you received may indicate that CmdStan was not installed correctly or was not found by cmdstanr. You may need to reinstall CmdStan and ensure that it is installed in the correct location on your computer.

### Reinstall CmdStan

Once you have installed CmdStan correctly, you can try reinstalling cmdstanr using the following command in RStudio:

-   `**install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))**`

**This will reinstall the cmdstanr** package and ensure that it is pointing to the correct CmdStan installation directory.

After reinstalling cmdstanr, you can try running the brm() function again and see if it works without any errors.

### Installing CmdStran

To install CmdStan on your computer, you need to follow these steps:

1.  Go to the CmdStan website: https://mc-stan.org/users/interfaces/cmdstan.

2.  Download the latest version of CmdStan that is compatible with your operating system (Windows, Mac, or Linux).

3.  Extract the downloaded file to a directory of your choice. By default, CmdStan installs in the home directory of your computer and creates a folder named "cmdstan-{version}". You should check that this folder exists and contains the necessary files.

4.  Open a terminal or command prompt and navigate to the directory where you extracted CmdStan.

5.  **Build the CmdStan executables** by running the following command:

-   `**make build**`

6.  Test the CmdStan installation by running the following command:

-   a`**make test**`

If the installation was successful, you should see a message indicating that all tests have passed.

7.  **Once you have installed CmdStan correctly, you can install the cmdstanr package in RStudio** and use it to run Bayesian models. You can install cmdstanr using the following command:

```{=html}
<!-- -->
```
a.  `**install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))**`

b.  Make sure to set the correct path to the CmdStan installation directory using the set_cmdstan_path() function before running the brm() function in RStudio.

### Installing *Cygwyn*

If you do not have Cygwin installed on your computer, you can download and install it from the official website: https://www.cygwin.com/. Make sure to install the latest version of Cygwin that is compatible with your operating system.

Once you have installed Cygwin, you should be able to locate the cygwin1.dll file in the Cygwin installation directory. By default, the cygwin1.dll file should be located in the bin subdirectory of the Cygwin installation directory.

If you have installed Cygwin in a non-default location or cannot locate the cygwin1.dll file, you may need to reinstall Cygwin or contact the Cygwin support team for assistance.

After you have located the cygwin1.dll file, follow the steps I mentioned in my previous response to delete all versions of the file except for the most recent version, and then try running the make build command again to build the CmdStan executables.

If you are installing Cygwin from the official website, you will be presented with a list of packages to install. The packages you need to install depend on the specific features and functionality you require.

### To build the CmdStan executables, you need to install the following packages:

-   `**make**`: this package provides the make command, which is used to build software from source code.

-   `**gcc-core and gcc-g++**`: these packages provide the GNU C and C++ compilers, which are used to compile and link the CmdStan source code.

-   `**mingw64-x86_64-gcc-g++**`: this package provides the *MinGW GCC C++ compiler*, which is used to compile and link the CmdStan Math library.

You may also need to install additional packages depending on the specific features and functionality you require. For example, if you plan to use the rstan package in RStudio, you will need to install the libcurl-devel package to enable HTTPS support.

Once you have selected the packages you need, you can proceed with the installation process. Make sure to select the "Install from Internet" option and choose a download site close to your location for faster download speeds.

After you have installed Cygwin and the necessary packages, you should be able to build the CmdStan executables by running the make build command in the CmdStan installation directory. If you encounter any issues, make sure to check the Cygwin documentation and support forums for assistance.
 